\documentclass{article}
\usepackage{soto-homework}

\newif\ifsolution
\input{solution}

\begin{document}

\chead{ENSP 330 Homework 1}
\hrule
\vspace{10pt}



% pedagogical goals
% - test back of envelope ability
% - show ability to convert units
% - find and interpret reliable sources of data
% - calculate overall efficiency from chain of efficiencies
% - understanding of energy conversion processes

% 2014 changes
% - give a more basic power and energy problem
% - give something that tests writing skills

You may submit this assignment handwritten.  Clearly state your
assumptions and cite your sources for any data not provided.


\problem{}

Worldwide energy use.  Given worldwide energy consumption at 474 EJ
   per year

\subproblem What is the volume of oil in barrels with this amount of chemical
        energy.

\subproblem How many kilograms of coal contain this much chemical energy?

\subproblem How many cubic feet of natural gas?

\subproblem If solar panels are 18\% efficient, how many square meters of
    solar panels in an area with 6.0 kWh per square meter per day?


\problem{} Estimate the equivalent power transfer when fueling a car.  Consider
a car with a 20 gallon gas tank.  Assume that it takes 3 minutes to fill
the tank.

\subproblem Calculate the rate at which gasoline flows (volume per time)

\subproblem Find the equivalent power transfer in units of watts

\subproblem How many hours during a disaster would this 20 gallons of
    gasoline power a 1 kW household generator with 40\% efficiency?



\problem{} How much gasoline is saved by the average car today versus 20 years
ago?  Use the CAFE standards for passenger cars.

\subproblem Find the average fuel efficiency for a U.S. vehicle in 2010 and
    in 1990.

\subproblem Assuming 15,000 miles per year and gasoline cost of 3.50 USD what
    is the fuel savings between the two for a single vehicle?

\subproblem What is the fuel savings in U.S. dollars for the entire nation?



\problem{} Energy Data.  Follow the link at the U.S. Energy Information
Administration and use the chart there to answer the following
questions.
% [http://www.eia.gov/totalenergy/data/annual/pecss_diagram.cfm](http://www.eia.gov/totalenergy/data/annual/pecss_diagram.cfm)


\subproblem Convert the 2011 electric energy use from quadrillion Btu into kWh.

\subproblem For each of the five supply sources shown, summarize their major
    uses in one sentence.

\subproblem What specific types of energy end uses do you imagine are included
    in each of the four demand sectors?


\problem{} Estimate to the nearest hour how much time you spent on this
assignment.



\end{document}
