\documentclass[12pt, oneside]{article}
\usepackage{graphicx}
\usepackage[nohead, margin=1.0in]{geometry}
%\usepackage{pythontex}

% this uses pythontex
% compile the document using > bash compile.sh

% set up whether we are printing assignment or solution
\newif\ifsolution
\solutiontrue
\solutionfalse

% page formatting
\pagestyle{empty}
\setlength{\parindent}{0pt}
\setlength{\parskip}{10pt}

% macros for formatting
\def\true{true}
\def\false{false}

% creates true false command with indented question
\newcommand{\tf}[1]
{
\vfill
\parbox[t]{0.25\textwidth}{\bf TRUE \hspace{0.1 in} FALSE }
\parbox[t]{0.75\textwidth}{#1 (1 pt)}
}

% creates grading table row macro
\newcommand{\tablerow}[2]
{\rule[-0.3cm]{0cm}{1cm}#1 & #2 & \\ \hline}

% creates heading for problems
\newcommand{\problem}[1]{{\bf Problem #1}}

% creates centered heading for sections
\newcommand{\chead}[1]
{\begin{center}\large\textbf{#1}\end{center}
\hrule
\vspace{10pt}}

% solution
\newcommand{\solution}[1]
{\ifsolution
Answer: {\it #1}
\else\fi}

% end macros



\begin{document}

% header block
\begin{center}
{\bf Midterm Exam}\\
{\bf ENSP 330}\\
{\bf Energy, Technology, and Society}\\
{\bf Thursday 16 Oct 2014}\\
\end{center}


% instructions to students
\noindent
{\bf Instructions}
\begin{itemize}
\item Do not begin until instructed.
\item You are allowed to use your cheat sheet only for the first 60
minutes.
\item After 60 minutes you may use your notes (written or electronic).
\item You may not use the internet with the exception of any notes you
keep on a cloud drive.
\item Clearly show your quantitative work and write your final answers
in the boxes provided.
\item In order to receive full credit, you must show your work with
units and justify your answers.  The correct answer without any work
will receive little or no credit.
\item Please bring to my attention anything that is unclear.
\item Write your name and signature that you understand these
instructions on the back page.
\item Use scientific notation where appropriate.
\item
\end{itemize}

% conversions
\noindent
{\bf Useful quantities}
\begin{itemize}
\item 1 nutritional calorie = 4186 joules
\item Gasoline contains 130 MJ per kilogram
\item Gasoline weighs 2.8 kilograms per gallon
\item kilo = $10^3$, mega = $10^6$, giga = $10^9$
\item Use 0.13 USD per kWh as the price of electricity
\end{itemize}

\vfill


%
% midterm content
%

\newpage
\chead{Quantitative}

% \begin{pycode}
% power_W = 200
% time_min = 15
%
% kilojoules = power_W * time_min * 60 / 1000.
% \end{pycode}

You are exercising on a StairMaster and the display says you expending
power a rate of 200 Watts.  How many joules of energy have you expended
after 15 minutes? (5 pts)

\framebox[2.5in]{\rule{0cm}{1.5cm}}\\

\solution{
Energy equals power multiplied by time.

%$$ 200W \cdot 15 min \cdot \frac{60 sec}{1 min} = \py{kilojoules} kJ $$
$$ 200W \cdot 15 min \cdot \frac{60 sec}{1 min} = 180 kJ $$
}


\vfill

If I ate a breakfast of waffles with maple syrup worth 500 nutritional
calories, how
long would it take me to use up those calories on the StairMaster
assuming you convert 100\% of your breakfast into kinetic energy? (5
pts)

\framebox[2.5in]{\rule{0cm}{1.5cm}}\\

\solution{
Here I assume that we are 100\% efficient in the conversion of food chemical
energy to kinetic energy.  Using energy equals power multiplied by time
and the rate of energy use we determined in c:

$$ 500 kcal = \frac{43 kcal}{15 min} \cdot X min $$
$$ X = 500 kcal \cdot  \frac{15 min}{43 kcal} = 174 min $$

Which is slightly less than 3 hours.
}

\vfill

\newpage
\chead{Quantitative}

Gasoline costs the military as much as 15 USD per gallon in some parts
of the world.  How much does this cost in USD/MJ?  For comparison,
electricity is about 0.028 USD per kWh. (5 pts)

\framebox[2.5in]{\rule{0cm}{1.5cm}}\\

\vfill

Assume a Tesla costs 80 thousand dollars to buy.  Model
S consumes 35 kWh to travel 100 miles.  Assume you keep the car for 10
years and drive 15,000 miles per year.  What is
the levelized cost per mile? For comparison, a Honda Fit is about 0.33
USD per mile.  (10 pts)

\framebox[2.5in]{\rule{0cm}{1.5cm}}\\


\vfill

\newpage
\chead{Short Answer}

Describe the important differences and similarities between how a coal
plant and a nuclear plant generate electricity. (5 pts)

\solution{Differences: The main difference is in the source of heat.  A
coal plant uses combustion of coal while a nuclear plant uses a chain
reaction of nuclear fission to create heat.

Similarities: The main
similarity is in the use of steam turbines to generate electricity
through a steam heat engine.}

\vfill
%\vspace{1 in}

Describe the important differences between a solar photovoltaic
electricity plant and a solar thermal electricity plant. (5 pts)

\solution{Solar PV electricity creates direct current electricity
without a  thermal cycle while solar thermal electricity uses technology
Rankine cycle steam turbines to generate AC electricity.}

\vfill
%\vspace{1 in}

Describe a few advantages and disadvantages of large, centralized
electricity production. (5 pts)

\solution{Advantages include economies of scale that make larger plants
more cost effective.  Disadvantages include large investments, large
effect from individual plant failures, and
transmission costs.}

\vfill
%\vspace{1 in}

What are the main uses in the U.S. economy for coal, oil, and natural gas? (5 pts)

\solution{Coal is mostly used for electricity production, oil is mostly
used for transportation, natural gas has many uses including
electricity, heating, and transportation.}

%\vspace{1 in}
\vfill

What is the original source of our food energy? (5 pts)

\solution{The chemical energy in our food was originally radiation
energy from the sun which was converted to chemical energy by plants.
Depending on your diet, you may also consume organisms that consumed
plants.  Bonus if you mention the nuclear energy from the sun.}

\vfill


\newpage
\chead{True or False?}

% \tf{A 100 MW solar photovoltaic electricity plant will likely produce
% more power in a year than a 100 MW coal electricity plant.}
% \solution{False}
%
\tf{Almost all of the energy from fossil fuels is converted to
electricity in a thermal power plant.}
\solution{False}

\tf{An energy bar contains less chemical energy than a gallon of
gasoline.}
\solution{True}

\tf{A large fraction of oil is used for electricity production in the
United States.}
\solution{False}

\tf{The BTU, the joule, and the Watt are all units of energy.}
\solution{False}

\tf{The kilowatt-hour (kWh) is a unit of power.}
\solution{False}

\tf{The second law of thermodynamics tells us that energy is conserved.}
\solution{False}

\tf{The energy intensity of an economy expresses how efficiently an
economy uses energy to create wealth per unit of GDP.}
\solution{True}

% \tf{The scale of your household daily electricity energy use is about 1 MWh.}
% \solution{False}
%
\tf{A battery converts chemical energy to electrical energy}
\solution{True}

\tf{Coal is a primary energy source.}
\solution{True}

\tf{Electricity is a primary energy source.}
\solution{False}


\newpage

% name and signature boxes
\makebox[1.0in][l]
{Name:}
\framebox[4.5in]{\rule{0cm}{1.5cm}}\\
\vspace{0.2cm}

\makebox[1.0in][l]
{Signature:}
\framebox[4.5in]{\rule{0cm}{1.5cm}}\\
\vspace{0.8cm}

By signing this you agree to all the instructions on the front page.

\vfill

% grading and scores table
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
\rule[-0.3cm]{0cm}{1cm}
Question & Points & Score \\
\hline
\tablerow{Quantitative}{25}
\tablerow{Short Answer}{15}
\tablerow{True/False}{10}
\tablerow{\bf{Total}}{50}
\end{tabular}
\end{center}
\end{document}

% do energy questions exist on the web?

% solar questions

% nuclear questions

% gdp questions

% fossil fuel questions

% thermo questions

% energy units questions

% estimate the capacity factor of an automobile

Convert from kWh to joules.

You have 1 kg of coal, how much hot water can you create?

