\documentclass[12pt, oneside]{article}
\usepackage{graphicx}
\usepackage[nohead, margin=1.0in]{geometry}


% set up whether we are printing assignment or solution
\newif\ifsolution
\solutiontrue
\solutionfalse

% page formatting
\pagestyle{empty}
\setlength{\parindent}{0pt}
\setlength{\parskip}{10pt}

% creates true false command with indented question
\newcommand{\tf}[1]
{
\vfill
\parbox[t]{0.25\textwidth}{\bf TRUE \hspace{0.1 in} FALSE }
\parbox[t]{0.75\textwidth}{#1 (2pt)}
}

% creates heading for problems
\newcommand{\problem}[1]{{\bf Problem #1}}

% creates centered heading for sections
\newcommand{\chead}[1]
{\begin{center}\large\textbf{#1}\end{center}
\hrule
\vspace{10pt}}

% solution
\newcommand{\solution}[1]
{\ifsolution
Answer: {\it #1}
\else\fi}

% end macros

\begin{document}

{\bf Quiz \hfill ENSP 330 \hfill 12 Nov 2013 }

% name and signature boxes
\makebox[1.0in][l]
{Name:}
\framebox[4.5in]{\rule{0cm}{1.5cm}}\\

\vfill

1) Name one method scientists use to infer temperature or climate
conditions from before
modern thermometers were in widespread use.

\solution{Tree rings, ice cores, pollen, coral, or others}

\vfill
2) The fraction of carbon dioxide in the atmosphere is closest to\\
a) 300 parts per million \\
b) 400 parts per million \\
c) 500 parts per million

\solution{b. as of 9 Nov 2013, Mauna Loa Observatory reads 394.57 ppm}

\vfill
3) In the Pacala and Socolow reading, what does BAU mean?

\solution{Business as usual.  Specifically, this is the amount of carbon
emissions if we do nothing different and continue economic growth.}

\vfill
4) If all the wedges recommended by Pacala and Socolow are successful what
will happen to the rate of our carbon emissions?

\solution{The goal of the wedges is to hold carbon emissions for the
next 50 years to the level they are now, approximately 7 GtC per year.}

\vfill
5) If all the wedges recommended by Pacala and Socolow are implemented
what will happen to the total amount of carbon in the atmosphere?

\solution{The goal is to prevent total carbon dioxide levels from
exceeding 500 ppm.}

\vfill




\end{document}

